Biometry of the invasive lionfish (*Pterois volitans*) in Playa del Carmen, Mexico and a review of length - weight parameters across the invasion range

### 2017-07-07 Manuscript.Rmd
Updated manuscript

---
title: "Manuscript"
output:
  html_notebook:
    code_folding: hide
  word_document: default
bibliography: references.bib
---

# Title page

**Title:** Biometry of the invasive lionfish (*Pterois volitans*) in Playa del Carmen, Mexico and a review of length - weight parameters across the invasion range

**Name of author:** Juan Carlos Villaseñor-Derbez

**Affiliation:** Bren School of Environmental Sciences and Management

**Full postal address:** University of California Santa Barbara, Santa Barbara,  CA 93106, USA

**Running title:** Length - weight parameters for *Pterois volitans*

@@

Juan Carlos Villaseñor-Derbez

4312 Bren Hall

Santa Barbara, CA

93106

jvillasenor@bren.ucsb.edu

+1 (207) 205 8435

@@

**Title**

@@

300 words

@@

lionfish, biometry, length-weight relationship, Mexico

___________________

@@

@@

pez león, biometría, relación lopngitud-peso, México

@@

```{r}

# source('./update_ref.R')
# update_ref()

```{r}
fish <- read_csv("./lionfish_data.csv", col_types = cols()) %>% 
  group_by(Id, Site, Location, Latitude, Longitude, Depth_m) %>% 
  summarize(Length = 10*mean(TotalLength_cm),
            Weight = mean(TotalWeigth_gr),
            Stomach = mean(StomachLength_cm)) %>% 
  ungroup()
```

@@

```{r}
# Load shapefiles
proj <- CRS("+proj=longlat +datum=NAD27")

@@

```{r, results = "asis"}
total <- data.frame(Location = "Total", Latitude = "", Longitude = "", N = 109, stringsAsFactors = F)

fish %>% 
  group_by(Location) %>% 
  summarize(Latitude = formatC(mean(Latitude), digits = 3, format = "f"),
            Longitude = formatC(mean(Longitude), digits = 3, format = "f"),
            N = n()) %>% 
  rbind(total) %>% 
  stargazer(type = "html", header = F, summary = F, add.lines = list(c("Total", "", "", 109)), rownames = F)
```

@@

```{r, results = "asis"}
fish %>% 
  select(Length, Weight, Stomach, Depth_m) %>% 
  as.data.frame() %>% 
  stargazer(header = F, type = "html", digits = 2)
```

```{r, fig.width = 5, fig.height = 7}
length_hist <- ggplot(data = fish, mapping = aes(x = Length)) +
  geom_density(fill = "gray", alpha = 0.5) +
  theme_bw() +
  ggExtra::removeGrid() +
  labs(x = "Total Length (mm)", y = "Density") +
  geom_text(label = "a)", x = 310, y = 0.007) +
  theme(text = element_text(family = "Times"))

@@

```{r}
model <- lm(log10(Weight) ~ log10(Length), data = fish)
```

```{r, results = "asis"}
TidyModel <- model%>%
  lmtest::coeftest(vcov = sandwich::vcovHC(., type = "HC1")) %>%
  broom::tidy()

@@

stargazer(model,
          header = F,
          se = list(robust_se),
          type = "html",
          t.auto = T,
          p.auto = T,
          intercept.bottom = F,
          single.row = T,
          report = "vcs*",
          omit.stat = c("rsq", "f"),
          add.lines = list(f_string),
          ci = T)
```

@@

```{r}
review <- read_csv("biometric_review.csv", col_types = cols())

knitr::kable(review, digits = 2, row.names = F, align = "l")
```

```{r}

@@

plot_fit_this_study <-  ggplot(data = fish, mapping = aes(x = Length, y = Weight)) +
  geom_line(data = predicted_this_study, mapping = aes(x = Length, y = Weight), size = 1) +
  geom_point(size = 2, alpha = 0.25, color = "black") +
  theme_bw() +
  ggExtra::removeGrid() +
  labs(x = "Total length (mm)", y = "Total Weight (g)") +
  theme(text = element_text(family = "Times")) +
  scale_color_brewer(palette = "Set3") +
  theme(text=element_text(family="Times"))

@@

```{r, fig.height = 9.5, fig.width = 6}
predicted_all <- data.frame(Length = seq(min(fish$Length), max(fish$Length), by = 1)) %>% 
  mutate(This = 10^a*Length^b, 
         APA = (10^review$a[1]) * ((Length/10)^review$b[1]),
         APF = (10^review$a[2]) * (Length^review$b[2]),
         APM = (10^review$a[3]) * (Length^review$b[3]),
         Sandel = review$a[4] * ((Length/10)^review$b[4]),
         Chin = (10^review$a[5]) * (Length^review$b[5]),
         Barbour = review$a[6] * Length^review$b[6],
         deLeon = review$a[7] * Length^review$b[7],
         FoggA = (10^review$a[8]) * (Length^review$b[8]),
         FoggM = (10^review$a[9]) * (Length^review$b[9]),
         FoggF = (10^review$a[10]) * (Length^review$b[10]),
         Edwards = review$a[11] * Length^review$b[11],
         Sabido = review$a[12] * ((Length/10)^review$b[12]),
         FishBase = review$a[13] * ((Length/10)^review$b[13])) %>%
  gather(Study_id, Weight, -Length) %>% 
  left_join(review, by ="Study_id") %>% 
  mutate(ThisStudy = ifelse(Study_id == "This", "Actual", "Other"))

plot_fit_all_studies <- ggplot(data = fish, mapping = aes(x = Length, y = Weight)) +
  geom_line(data = predicted_all, mapping = aes(x = Length, y = Weight, linetype = Gender, color = Study), size = 1) +
  geom_point(size = 2, alpha = 0.25, color = "black") +
  theme_bw() +
  ggExtra::removeGrid() +
  labs(x = "Total length (mm)", y = "Total Weight (g)") +
  theme(text = element_text(family = "Times"),
        legend.justification=c(0,1),
        legend.position=c(0.01, 0.99)) +
  guides(color = guide_legend(title=NULL)) +
  scale_color_brewer(palette = "Paired")

plot_fit_all_studies_log <- ggplot(data = fish, mapping = aes(x = Length, y = log10(Weight))) +
  geom_line(data = predicted_all, mapping = aes(x = Length, y = log10(Weight), linetype = Gender, color = Study), size = 1) +
  geom_point(size = 2, alpha = 0.25, color = "black") +
  theme_bw() +
  ggExtra::removeGrid() +
  labs(x = "Total length (mm)", y = "log10(TW)") +
  theme(text = element_text(family = "Times"),
        legend.position = "None")+
  scale_x_continuous(expand=c(0,0))+
  scale_y_continuous(expand=c(0,0))

@@

@@

# Raw data

>>>>

---
title: "Biometry of the invasive lionfish (*Pterois volitans*) in Playa del Carmen, Mexico and a review of length - weight parameters across the invasion range"
header-includes:
- \usepackage{setspace}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
output:
  pdf_document:
    fig_caption: yes
fontsize: 12pt
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

# source('./update_ref.R')
# update_ref()
```

```{r, load packages}

Juan Carlos Villaseñor-Derbez^1^\

^1^ Bren School of Environmental Sciences and Management, University of California Santa Barbara\

**Full postal address:**
4312 Bren Hall, Santa Barbara, CA, 93106\

**Running title:** Length - weight parameters for *Pterois volitans* \

@@

Juan Carlos Villaseñor-Derbez   
4312 Bren Hall  
Santa Barbara, CA   
93106   
jvillasenor@bren.ucsb.edu   
+1 (207) 205 8435

@@

**Title of the contribution:** Biometry of the invasive lionfish (*Pterois volitans*) in Playa del Carmen, Mexico and a review of length - weight parameters across the invasion range\

@@

300 words\

@@

lionfish, biometry, length-weight relationship, Mexico\

@@

300 palabras\

@@

pez león, biometría, relación lopngitud-peso, México\

@@

```{r load fish data, include = F}
fish <- read_csv("./lionfish_data.csv", col_types = cols()) %>% 
  group_by(Id, Site, Location, Latitude, Longitude, Depth_m) %>% 
  summarize(Length = 10*mean(TotalLength_cm),
            Weight = mean(TotalWeigth_gr),
            Stomach = mean(StomachLength_cm)) %>% 
  ungroup()
```

@@

```{r create map, include = F}
# Load shapefiles
proj <- CRS("+proj=longlat +datum=NAD27")

@@

```{r table of locations, include = F}
total <- data.frame(Location = "Total", Latitude = "", Longitude = "", minD = min(fish$Depth_m), maxD = max(fish$Depth_m), meanD = formatC(mean(fish$Depth_m), digits = 1, format = "f"), n = 109, stringsAsFactors = F)

location_table <- fish %>% 
  group_by(Location) %>% 
  summarize(Latitude = formatC(mean(Latitude), digits = 3, format = "f"),
            Longitude = formatC(mean(Longitude), digits = 3, format = "f"),
            minD = formatC(min(Depth_m), digits = 1, format = "f"),
            maxD = formatC(max(Depth_m), digits = 1, format = "f"),
            meanD = formatC(mean(Depth_m), digits = 1, format = "f"),
            n = n()) %>% 
  rbind(total) %>% 
  knitr::kable(row.names = F, caption = "Coordinates, minimum, maximum and mean depth (m), and number of samples for each location. n = sample size.", col.names = c("Location", "Latitude", "Longitude", "Minimum Depth (m)", "Maximum Depth (m)", "Mean Depth (m)", "n"))

location_table
```

@@

```{r summary table, include = F}
summary_table <- fish %>% 
  select(Length, Weight) %>% 
  as.data.frame() %>% 
  stargazer(header = F, type = "latex", digits = 2, title = "Summary statistics of Length and Weight of sampled organisms.")
```

```{r histogram figures, include = F}
length_hist <- ggplot(data = fish, mapping = aes(x = Length)) +
  geom_density(fill = "gray", alpha = 0.5) +
  theme_bw() +
  ggExtra::removeGrid() +
  labs(x = "Total Length (mm)", y = "Density") +
  geom_text(label = "a)", x = 310, y = 0.007) +
  theme(text = element_text(family = "Times"))

@@

```{r model fitting}
model <- lm(log10(Weight) ~ log10(Length), data = fish)
```

```{r regression table, include = F}
TidyModel <- model%>%
  lmtest::coeftest(vcov = sandwich::vcovHC(., type = "HC1")) %>%
  broom::tidy()

@@

regression_table <- stargazer(model,
          header = F,
          se = list(robust_se),
          type = "latex",
          t.auto = T,
          p.auto = T,
          intercept.bottom = F,
          single.row = T,
          report = "vcs*",
          omit.stat = c("rsq", "f"),
          add.lines = list(f_string),
          ci = T)
```
\clearpage

@@

```{r, include = F}
review <- read_csv("biometric_review.csv", col_types = cols())

review_table <- review %>% 
  select(Reference, n = N, a, c, b, r2, Gender, minD, maxD, Spear) %>% 
  knitr::kable(digits = 4, row.names = F, caption = "Allometric growth parameters for eight published papers, Fishbase (Froese and Pauly 2016), and this study. All parameters have been adjusted to convert from millimiters to grams. n = Sample size, a = scaling parameter for eq. 1, c = y-intercept for eq. 3, b = exponent or slope for eq. 1 or eq. 2, respectively.")

review_table
```

```{r scatter plot, include = F}

@@

plot_fit_this_study <-  ggplot(data = fish, mapping = aes(x = Length, y = Weight)) +
  geom_line(data = predicted_this_study, mapping = aes(x = Length, y = Weight), size = 1) +
  geom_point(size = 2, alpha = 0.25, color = "black") +
  theme_bw() +
  labs(x = "Total length (mm)", y = "Total Weight (g)") +
  theme(text = element_text(family = "Times")) +
  scale_color_brewer(palette = "Set3") +
  theme(text=element_text(family="Times"))

@@

```{r review plots, include = F}
predicted_all <- data.frame(Length = seq(min(fish$Length), max(fish$Length), by = 1)) %>% 
  mutate(This = review$a[14] * (Length^review$b[14]), 
         APA = review$a[1] * (Length^review$b[1]),
         APF = review$a[2] * (Length^review$b[2]),
         APM = review$a[3] * (Length^review$b[3]),
         Sandel = review$a[4] * (Length^review$b[4]),
         Chin = review$a[5] * (Length^review$b[4]),
         Barbour = review$a[6] * (Length^review$b[6]),
         deLeon = review$a[7] * (Length^review$b[7]),
         FoggA = review$a[8] * (Length^review$b[8]),
         FoggM = review$a[9] * (Length^review$b[9]),
         FoggF = review$a[10] * (Length^review$b[10]),
         Edwards = review$a[11] * (Length^review$b[11]),
         Sabido = review$a[12] * (Length^review$b[12]),
         FishBase = review$a[13] * (Length^review$b[13])) %>%
  gather(Study_id, Weight, -Length) %>% 
  left_join(review, by ="Study_id") %>% 
  mutate(ThisStudy = ifelse(Study_id == "This", "Actual", "Other"))

plot_fit_all_studies <- ggplot(data = fish, mapping = aes(x = Length, y = Weight)) +
  geom_line(data = predicted_all, mapping = aes(x = Length, y = Weight, linetype = Gender, color = Study), size = 1) +
  theme_bw() +
  labs(x = "Total length (mm)", y = "Total Weight (g)") +
  theme(text = element_text(family = "Times"),
        legend.justification=c(0,1),
        legend.position=c(0.01, 0.99),
        legend.background = element_rect(fill = "transparent", size=.5, linetype="dotted")) +
  guides(color = guide_legend(title=NULL)) +
  scale_color_brewer(palette = "Paired")

plot_fit_all_studies_log <- ggplot(data = fish, mapping = aes(x = Length, y = log10(Weight))) +
  geom_line(data = predicted_all, mapping = aes(x = Length, y = log10(Weight), linetype = Gender, color = Study), size = 1) +
  theme_bw() +
  labs(x = "Total length (mm)", y = "log10(TW)") +
  theme(text = element_text(family = "Times"),
        legend.position = "None")+
  scale_x_continuous(expand=c(0,0))+
  scale_y_continuous(expand=c(0,0))

@@

# Tables

## Location table

```{r}
location_table
```

## Summary table

```{r, results = "asis"}
cat(summary_table)
```

## Regression table

```{r, results = "asis"}
cat(regression_table)
```

## Review table

```{r}
review_table
```

# Figures

## Map

```{r}
tmap_arrange(pen_map, study_area, outer.margins = 0.005, ncol = 2)
```

## Histograms

```{r, fig.width = 5, fig.height = 7}
gridExtra::grid.arrange(length_hist, weight_hist)
```

## Scatter plot

```{r}
plot_fit_this_study
```

## Review plots

```{r, fig.height = 9.5, fig.width = 6}
gridExtra::grid.arrange(plot_fit_all_studies, plot_fit_all_studies_log)
```

@@
